Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2019-01-15 paper/paper.tex
Minor changes

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with
time in the LCR (e.g. \Fig{} \ref{fig:2}a) within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes (blue bars in \Fig{}
\ref{fig:2}b) of different salts were obtained by converting the
conductivity-time profiles using calibration curves (Supplementary
Section \ref{SI-sec:measure}). The theoretical diffusive fluxes
through PCTE are obtained based on the assumption that the membrane
consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:
\begin{equation}
  \label{eq:j-pcte}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi r_{\mathrm{PCTE}}^{2} N_{\mathrm{p}} D_{\mathrm{\pm}} \Delta c}{L \tau}
\end{equation}

where $r_{\mathrm{PCTE}}$ is the PCTE pore radius, $N_{\mathrm{p}}$ is
numbers of pore per area, $D_{\mathrm{\pm}}$ is the salt diffusivity
(average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $L$ is the thickness of the PCTE membrane.
By using the salt diffusivity values (Supplementary Table
\ref{SI-tab:diff}), the calculated diffusive fluxes (red bars in
\Fig{} \ref{fig:2}b) show reasonable agreement with the experimental
values.

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene's
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:

\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi)
  =
  - N_{\mathrm{A}} e \sum_{i} c_{i} z_{i}
\end{equation}

where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
individual materials in the system, including water and the internal
Stern layer\cite{Tian_2017}, and
$\varepsilon_{0}$ is the vacuum permittivity. When applying
$V_{\mathrm{G}}$ to graphene adjacent to the electrolyte solution,
charges (electrons or holes) are induced in graphene; the
electroneutrality of the entire system suggests:
\begin{equation}
  \label{eq:electro-neutral}
  \sigma_{\mathrm{G}} S_{\mathrm{G}} + \sum_{i} \int_{\Omega} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega= 0
\end{equation}
where $\sigma_{\mathrm{G}}$ is the charge density in graphene,
$S_{\mathrm{G}}$ is the total area of graphene, and $\Omega$
corresponds to the entire volumetric domain of electrolyte
solution. Note that the graphene surface potential,
$\psi_{\mathrm{G}}$, is not equivalent to $V_{\mathrm{G}}$ due to
change of graphene's work function upon charging, known as the
graphene quantum capacitance effect\cite{Xia_2009} (\Fig{}
\ref{fig:3}a) following:

\begin{equation}
  \label{eq:Vg}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}
\end{equation}

where
$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be 1.4$\times$10$^{11}$ \textit{e}$\cdot$cm$^{-2}$, corresponding
to a work function difference of 48 meV from graphene's Dirac point
(details see Supplementary Section \ref{SI-sec:charge-dens}), much
smaller than the $\Delta \phi_{\mathrm{G}}$ induced by gating.
Therefore we adapt the simplification that the charge neutrality point
of graphene coincides with the Dirac point at $V_{\mathrm{G}}$ =
0. The elementary electronic properties of graphene give:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay
between graphene quantum capacitance and the EDL significantly reduces
the graphene surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the system here is that the
surface charge density can be precisely determined and controlled,
rather than being treated as a fitting parameter as in most of the 2D
nanopore studies \cite{Rollings_2016}. We anticipate that the
gate-controlled ionic transport can also be achieved by replacing
graphene with other types of conducting 2D materials, although the
operational gate voltage window required to achieve the same order of
salt rejection may be narrower than graphene due to larger quantum
capacitances \cite{Tian_2016}.

>>>>

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with
time in the LCR (e.g. \Fig{} \ref{fig:2}a) within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes (blue bars in \Fig{}
\ref{fig:2}b) of different salts were obtained by converting the
conductivity-time profiles using calibration curves (Supplementary
Section \ref{SI-sec:measure}). The theoretical diffusive fluxes
through PCTE are obtained based on the assumption that the membrane
consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:
\begin{equation}
  \label{eq:j-pcte}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi r_{\mathrm{PCTE}}^{2} N_{\mathrm{p}} D_{\mathrm{\pm}} \Delta c}{L \tau}
\end{equation}
where $r_{\mathrm{PCTE}}$ is the PCTE pore radius, $N_{\mathrm{p}}$ is
numbers of pore per area, $D_{\mathrm{\pm}}$ is the salt diffusivity
(average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $L$ is the thickness of the PCTE membrane.
By using the salt diffusivity values (Supplementary Table
\ref{SI-tab:diff}), the calculated diffusive fluxes (red bars in
\Fig{} \ref{fig:2}b) show reasonable agreement with the experimental
values.

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene's
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:
\begin{equation}
  \label{eq:poisson}
  \nabla \cdot (\varepsilon_{\mathrm{m}} \varepsilon_{0} \nabla \psi)
  =
  - N_{\mathrm{A}} e \sum_{i} c_{i} z_{i}
\end{equation}
where $\varepsilon_{\mathrm{m}}$ is the relative permittivity of
individual materials in the system, including water and the internal
Stern layer\cite{Tian_2017}, and
$\varepsilon_{0}$ is the vacuum permittivity. When applying
$V_{\mathrm{G}}$ to graphene adjacent to the electrolyte solution,
charges (electrons or holes) are induced in graphene; the
electroneutrality of the entire system suggests:
\begin{equation}
  \label{eq:electro-neutral}
  \sigma_{\mathrm{G}} S_{\mathrm{G}} + \sum_{i} \int_{\Omega} z_{i} c_{i} N_{\mathrm{A}} e \mathrm{d}^{3} \Omega= 0
\end{equation}
where $\sigma_{\mathrm{G}}$ is the charge density in graphene,
$S_{\mathrm{G}}$ is the total area of graphene, and $\Omega$
corresponds to the entire volumetric domain of electrolyte
solution. Note that the graphene surface potential,
$\psi_{\mathrm{G}}$, is not equivalent to $V_{\mathrm{G}}$ due to
change of graphene's work function upon charging, known as the
graphene quantum capacitance effect\cite{Xia_2009} (\Fig{}
\ref{fig:3}a) following:
\begin{equation}
  \label{eq:Vg}
  V_{\mathrm{G}} = \Delta \phi_{\mathrm{G}} + \psi_{\mathrm{G}}
\end{equation}
where
$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be 1.4$\times$10$^{11}$ \textit{e}$\cdot$cm$^{-2}$, corresponding
to a work function difference of 48 meV from graphene's Dirac point
(details see Supplementary Section \ref{SI-sec:charge-dens}), much
smaller than the $\Delta \phi_{\mathrm{G}}$ induced by gating.
Therefore we adapt the simplification that the charge neutrality point
of graphene coincides with the Dirac point at $V_{\mathrm{G}}$ =
0. The elementary electronic properties of graphene give:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay
between graphene quantum capacitance and the EDL significantly reduces
the graphene surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the system here is that the
surface charge density can be precisely determined and controlled,
rather than being treated as a fitting parameter as in most of the 2D
nanopore studies \cite{Rollings_2016}. We anticipate that the
gate-controlled ionic transport can also be achieved by replacing
graphene with other types of conducting 2D materials, although the
operational gate voltage window required to achieve the same order of
salt rejection may be narrower than graphene due to the larger quantum
capacitances \cite{Tian_2016}.
